An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-05 implementation.tex
Updates from Overleaf
fixed third property

prop spawning : 
  Forall x . Forall y . 
    spawn(x,y) ->  (P main(x) | Exists z . (P main(z) & leadsto(z,y))) 
    where leadsto(z,y) := 
      spawn(z,y) | @ leadsto(z,y) | Exists x . (@leadsto(z,x) & spawn(x,y))
\end{lstlisting}
\caption{Three properties stated in \dejavu's logic}
\label{fig:properties}
\end{figure}
\end{center}

@@

The third property, \idsl{spawning}, expresses a property
about threads being spawned in an operating system. We want to 
ensure that all threads are spawned by the main program or are spawned by a thread that transitively can be traced back to a thread being spawned by the main program. The events are
\idsl{main(x)} (the main program spawns thread \idsl{x}) and
\idsl{spawn(x,y)} (thread \idsl{x} spawns thread \idsl{y}).
For this we need to compute a transitive closure of spawning relationships, here expressed with the rule \idsl{leadsto(z,y)}.

@@

We evaluated the three properties in Figure 
\ref{fig:properties} on a collection of traces. Table \ref{tab:evaluation} shows the  
analysis time (excluding time to compile the generated 
monitor) for different traces (format is `trace length : seconds'). The telemetry traces will repeatedly open hundreds of channels and send telemetry over them, and the spawning traces will spawn a new thread for each event in the trace. The transitive closure of the spawning property needs to process these, therefore the more extensive processing time.

@@

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
Property   & Trace 1 & Trace 2 & Trace 3 \\
\hline
telemetry1 & 1,200,001\,:\,3.3s & 5,200,001\,:\,7.1s & 10,200,001\,:\,12.8s \\
\hline
telemetry2 & 1,200,001\,:\,4.4s & 5,200,001\,:\,9.9s & 10,200,001\,:\,17.9s \\
\hline
spawning   & 4,950\,:\,9.4s & 10,000\,:\,88.4s & 19,900\,:\,488.8s \\
\hline
\end{tabular}
\caption{Evaluation - trace lengths and analysis time in seconds}
\label{tab:evaluation}
\end{table}

>>>>

prop spawning :
  Forall x . Forall y . Forall d . report(y,x,d) -> spawned(x,y) where
    spawned(x,y) := @ spawned(x,y) | spawn(x,y) | 
                    Exists z . (@spawned(x,z) & spawn(z,y))
\end{lstlisting}
\caption{Three properties stated in \dejavu's logic}
\label{fig:properties}
\end{figure}
\end{center}

@@

The third property, \idsl{spawning}, expresses a property
about threads being spawned in an operating system. We want to 
ensure that when a thread \idsl{y} reports some data \idsl{d} 
back to another thread \idsl{x}, then thread \idsl{y} has been spawned by thread \idsl{x} either directly, or transitively 
via a sequence of spawn events. The events are
\idsl{spawn(x,y)} (thread \idsl{x} spawns thread \idsl{y})
and \idsl{report(y,x,d)} (thread \idsl{y} reports data \idsl{d}
back to thread \idsl{x}).
For this we need to compute a transitive closure of spawning relationships, here expressed with the rule \idsl{spawning(x,y)}.

@@

We evaluated the three properties in Figure 
\ref{fig:properties} on a collection of traces. Table \ref{tab:evaluation} shows the  
analysis time (excluding time to compile the generated 
monitor) for different traces (format is `trace length : seconds'). The telemetry traces will repeatedly open hundreds of channels and send telemetry over them. 
%
The spawning traces consist of 50\% spawning events and 50\% reporting events. The spawning property  requires larger processing time compared to the telemetry properties since more data (the transitive closure) has to be computed and stored.

@@

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
Property   & Trace 1 & Trace 2 & Trace 3 \\
\hline
telemetry1 & 1,200,001\,:\,3.3s & 5,200,001\,:\,7.1s & 10,200,001\,:\,12.8s \\
\hline
telemetry2 & 1,200,001\,:\,4.4s & 5,200,001\,:\,9.9s & 10,200,001\,:\,17.9s \\
\hline
spawning   & 9,899\,:\,42.7s & 19,999\,:\,127.6s & 39,799\,:\,572.4s \\
\hline
\end{tabular}
\caption{Evaluation - trace lengths and analysis time in seconds}
\label{tab:evaluation}
\end{table}
